Black hole spectroscopy using a complete gravitational wave signal from a binary black coalescence

### 2020-12-07 papers/intro_paper/intro_paper.tex
updated 2 figures
updated paper draft

\title{Black hole symphony orchestra: Spectroscopy using gravitational wave signals from binary black hole coalescences}

@@

We additionally investigate the effects of erroneously assuming that an underlying modified GR signal can be well-described using GR. We do this by estimating the parameters of our modified GR signals using the GR waveform model $\SEOB$ instead of the parametrised $\pSEOB$. In such cases, we run the risk of biased parameter estimates due to an incomplete understanding of the underlying signal. The resulting posterior probability distributions are shown in the right panels of Fig.~\ref{simulated_signal_nonGR} by the pink (GW150914) and grey (GW190521) curves. The results are interesting and distinctly different for the two events. For the GW150914-like modified GR signal, the measurements of $(\fgr{220}, \taugr{220})$ (Fig.~\ref{simulated_signal_nonGR} top right panel) are consistent with the $(\fgr{220}, \taugr{220})$ measurements for a signal with no deviations (Fig.~\ref{simulated_signal_GR} top right panel). In other words, if the actual signal had deviations as large as the 50 \% of the GR prediction, the analysis with $\SEOB$ would likely have reported \emph{no} deviation from the GR prediction. However in the case of the GW190521-like modified GR signal, a simple GR analysis of the modified GR signal would have yielded measurements distinctly different from either of the two parametrised estimates: with and without deviations. The fact that the GW190521-like signal has a much lower SNR than GW150914 might be a possible reason for the the measurement of the final quantities to be more susceptible to noise. A more detailed comparison of the other parameters, like the masses and spins, between an $pSEOB$ and an $SEOB$ measurement of a modified GR signal is shown in the appendix (Fig.~\ref{fig:gr_ngr_comparison}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
	\includegraphics[width=0.5\textwidth]{figures/modGR_waveforms_amplitudephase.png}
	\caption{\textcolor{red}{FINAL RESULT} Top panel: The `+'--polarisation of the gravitational waveform $h_+(t)$ from a GW150914-like event where the post-merger is described by GR (solid orange lines), i.e., $\df{220} = \dtau{220} = 0$, and where the merger-ringdown is modified (dashed grey lines), i.e., $\df{220} = \dtau{220} = 0.5$. Bottom panel: Comparison of the evolution of the amplitude (left) and phase (right) for the GR and modified GR signal.}
	\label{fig:nongr_waveform}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
	\includegraphics[width=0.5\textwidth]{figures/S190521g_swinjs.png}
	\caption{\textcolor{red}{FINAL RESULT} 90 \% credible level on the posterior probability distribution of the frequency and damping time of $(2,\pm 2)$ mode, $(\fgr{220}, \taugr{220}$ from simulations of an \texttt{NRSur7dq4} GW signal with parameters similar to the GW event, GW190521, in Gaussian noise (grey dot-dashed lines) and real interferometric noise (pink dot dashed lines). The GR prediction for the frequency and damping time is indicated by the black cross. While the Gaussian noise simulations are consistent with the prediction, at least 3 or the 5 real noise simulation are not. For comparison, we also plot the 90 \% credible level for the actual event, GW190521.}
	\label{fig:21g_systematics}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{Correlations between GR and non-GR parameters}\label{sec:correlations}
\abhi{Do we want a separate appendix to discuss Fig.~\ref{fig:gr_ngr_comparison}? Or is the reference in the main text currently, sufficient?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% modified GR signal: GR vs nonGR recovery comparison
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[h!]
	\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p5_gr_ngr_m1m2.png}\includegraphics[width=0.5\textwidth]{figures/GW190521_simulated_signal_0p5_gr_ngr_m1m2.png}
	\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p5_gr_ngr_a1za2z.png}\includegraphics[width=0.5\textwidth]{figures/GW190521_simulated_signal_0p5_gr_ngr_a1za2z.png}	
	\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p5_gr_ngr_fgrtaugr.png}\includegraphics[width=0.5\textwidth]{figures/GW190521_simulated_signal_0p5_gr_ngr_fgrtaugr.png}
	\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p5_gr_ngr_Mfaf.png}\includegraphics[width=0.5\textwidth]{figures/GW190521_simulated_signal_0p5_gr_ngr_Mfaf.png}
	\caption{\textcolor{red}{FINAL RESULT} Comparison of the recovered parameters when the underlying signal is assumed to be a GR signal or a modified GR signal. In both cases, the actual underlying signal is a modified GR signal with parameters similar to GW150914 (left panels) and GW190521 (right panels) respectively. The initial parameters are given in Table~\ref{tab:injection_values}, with the QNM parameters defined by $\df{220} = \dtau{220} = 0.5$. For the GW150914 (GW190521) contours, the $SEOB$ and $pSEOB$ recoveries are indicated by blue (red) and pink (grey) curves respectively. The panels (from top to bottom) show the recoveries in (detector-frame) initial masses (first row), z-components of dimensionless initial spins (second row), GR predictions of frequency and damping time (third row) and the remnant mass and spin predictions ($M_f, a_f$) from the frequency and damping time (obtained by inverting the Berti fits) CITE}
	\label{fig:gr_ngr_comparison}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\title{Black hole spectroscopy using a complete gravitational wave signal from a binary black coalescence}

@@

We additionally investigate the effects of erroneously assuming that an underlying modified GR signal can be well-described using GR. We do this by estimating the parameters of our modified GR signals using the GR waveform model $\SEOB$ instead of the parametrised $\pSEOB$. In such cases, we run the risk of biased parameter estimates due to an incomplete understanding of the underlying signal. The resulting posterior probability distributions are shown in the right panels of Fig.~\ref{simulated_signal_nonGR} by the pink (GW150914) and grey (GW190521) curves. The results are interesting and distinctly different for the two events. For the GW150914-like modified GR signal, the measurements of $(\fgr{220}, \taugr{220})$ (Fig.~\ref{simulated_signal_nonGR} top right panel) are consistent with the $(\fgr{220}, \taugr{220})$ measurements for a signal with no deviations (Fig.~\ref{simulated_signal_GR} top right panel). In other words, if the actual signal had deviations as large as the 50 \% of the GR prediction, the analysis with $\SEOB$ would likely have reported \emph{no} deviation from the GR prediction. However in the case of the GW190521-like modified GR signal, a simple GR analysis of the modified GR signal would have yielded measurements distinctly different from either of the two parametrised estimates: with and without deviations. The fact that the GW190521-like signal has a much lower SNR than GW150914 might be a possible reason for the the measurement of the final quantities to be more susceptible to noise. A more detailed comparison of the other parameters, like the masses and spins, between an $pSEOB$ and an $SEOB$ measurement of a modified GR signal is shown in Fig.~\ref{fig:gr_ngr_comparison}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
	\includegraphics[width=0.5\textwidth]{figures/modGR_waveforms_amplitudephase.png}
	\caption{\textcolor{red}{FINAL RESULT} Top panel: The `+'--polarisation of the gravitational waveform $h_+(t)$ from a GW150914-like event where the post-merger is described by GR (solid orange lines), i.e., $\df{220} = \dtau{220} = 0$, and where the merger-ringdown is modified (dashed grey lines), i.e., $\df{220} = \dtau{220} = 0.5$. Bottom panel: Comparison of the evolution of the amplitude (left) and instantaneous frequency (right) for the GR and modified GR signal.}
	\label{fig:nongr_waveform}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% modified GR signal: GR vs nonGR recovery comparison
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}%[h!]
	\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p5_gr_ngr_m1m2.png}\includegraphics[width=0.5\textwidth]{figures/GW190521_simulated_signal_0p5_gr_ngr_m1m2.png}
	\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p5_gr_ngr_a1za2z.png}\includegraphics[width=0.5\textwidth]{figures/GW190521_simulated_signal_0p5_gr_ngr_a1za2z.png}	
	\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p5_gr_ngr_fgrtaugr.png}\includegraphics[width=0.5\textwidth]{figures/GW190521_simulated_signal_0p5_gr_ngr_fgrtaugr.png}
	\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p5_gr_ngr_Mfaf.png}\includegraphics[width=0.5\textwidth]{figures/GW190521_simulated_signal_0p5_gr_ngr_Mfaf.png}
	\caption{\textcolor{red}{FINAL RESULT} Comparison of the recovered parameters when the underlying signal is assumed to be a GR signal or a modified GR signal. In both cases, the actual underlying signal is a modified GR signal with parameters similar to GW150914 (left panels) and GW190521 (right panels) respectively. The initial parameters are given in Table~\ref{tab:injection_values}, with the QNM parameters defined by $\df{220} = \dtau{220} = 0.5$. For the GW150914 (GW190521) contours, the $SEOB$ and $pSEOB$ recoveries are indicated by blue (red) and pink (grey) curves respectively. The panels (from top to bottom) show the recoveries in (detector-frame) initial masses (first row), z-components of dimensionless initial spins (second row), GR predictions of frequency and damping time (third row) and the remnant mass and spin predictions ($M_f, a_f$) from the frequency and damping time (obtained by inverting the Berti fits) CITE}
	\label{fig:gr_ngr_comparison}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\begin{center}
	\includegraphics[width=0.5\textwidth]{figures/S190521g_swinjs.png}
	\caption{\textcolor{red}{FINAL RESULT} 90 \% credible level on the posterior probability distribution of the frequency and damping time of $(2,\pm 2)$ mode, $(\fgr{220}, \taugr{220}$ from simulations of an \texttt{NRSur7dq4} GW signal with parameters similar to the GW event, GW190521, in Gaussian noise (grey dot-dashed lines) and real interferometric noise (pink dot dashed lines). The GR prediction for the frequency and damping time is indicated by the black cross. While the Gaussian noise simulations are consistent with the prediction, at least 3 or the 5 real noise simulation are not. For comparison, we also plot the 90 \% credible level for the actual event, GW190521.}
	\label{fig:21g_systematics}
\end{center}	
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@
